\LaTeX\ Author Guidelines for ICCV Proceedings

### 2019-03-27 chapters/3_algorithm.tex
Minor changes

\section{Unified Graph Agglomerative Clustering}
In this section we introduce the Unified Graph Agglomerative Clustering Algorithm (UGACA) that represents a simple generalized formalization of many agglomerative graph clustering algorithms. \UPDATE{It can be used to describe both unsigned clustering algorithms ingesting positive cluster similarities and signed clustering algorithms using attractive and repulsive cues.}\\
First, we define the graph notation and the concept of cannot-link constraints. Then, we introduce the generalized algorithm and show \UPDATE{how it can be used to introduce several new agglomerative clustering algorithms for graphs with signed weights.}

@@

Given an initial clustering $\Pi$ where each node of the original graph $\mathcal{G}$ represents its own cluster, the algorithm starts by initializing a contracted graph $\mathcal{G}'$ (see Definition \ref{def:contr_graph}) that will represent the clustering $\Pi$ at each iteration of the agglomeration. The algorithm then sorts all edges $e\in E$ by their absolute weight $|W(e)|$ in descending order in a priority queue, so that the most attractive and the most repulsive edges are processed first. It iteratively pops one edge $e_{uv}$ from priority queue and, depending on the value of the priority $\tilde{w}$ associated to $e_{uv}$, does the following. If the edge is attractive and $\tilde{w}>0$, then merge the connected nodes and perform an edge contraction of $e_{uv}$ in $\mathcal{G}'$ (see Sec. \ref{sec:contr_graph_and_CLC} and Fig. \ref{fig:edge_contraction_and_contr_graph}), provided that $e_{uv}$ was not previously flagged as a cannot-link constraint; for every new pair of double edges in $\mathcal{G}'$ update \UPDATE{their priorities} according to one of the update rules listed in Table \ref{tab:linkage-criteria} and update their cannot-link relationships (see example in Fig. \ref{fig:algorithm_with_without_CLC}b from step 3 to 4). If, on the other hand, the edge $e_{uv}$ is repulsive and $\tilde{w}\leq 0$, then flag it as a cannot-link constraint if and only if the algorithm input-option \texttt{addCannotLink} is true and mutual exclusion relationships should be enforced.

In the Supplementary material we provide a more detailed version of the algorithm implementation (Sec. \ref{sec:detailed_algorithm}) and we also comment on its theoretical complexity (Sec. \ref{sec:complexity}).

 In the special case of an unsigned graph %$\mathcal{G}(V,E,W:E\rightarrow \mathbb{R}^+)$ 
 presenting strictly positive similarities between nodes, the algorithm then returns only one single cluster as an output (see Fig. \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}a}), \UPDATE{together with a merging tree representing the merging order and defining an hierarchy of clusters} \TODO{use def of merging process of contracted graph to define merging tree?}.

@@

\subsection{Update rules and associated algorithms}
\begin{itemize}
\item In this section we will list some of the algorithms that can be described using the generalized algorithm we presented and we will introduce some new ones.
\item In Table \ref{tab:linkage-criteria} we collected the update rules that we tested and compared in this article. Later we will mention more of them that were proposed in the literature.
\item In the special case of unsigned graphs (Table \ref{tab:linkage-criteria} column 1), the UGAC algorithm becomes simply a version of hierarchical clustering with different linkage criteria depending on the chosen update rule. It outputs one single cluster and it also defines an hierarchy of clusters during the merging process that can be associated to \UPDATE{a merging tree/  the hierarchical clustering dendrogram}... \TODO{Problem: sum does not trivially define an ultra-metric-distance...}  
\item In the literature there have been already several methods using cannot-link constraints for agglomerative clustering algorithms \TODO{repetition from related work}: for example, \cite{malmberg2011generalized} used initially fixed mutual exclusion relationships, whereas \cite{levinkov2017comparative} introduced them dynamically in an edge contraction algorithm for signed graphs, named Greedy Fixation (Table \ref{tab:linkage-criteria} row 1, \emph{sum}). At the same time, they also proposed a version of it without cannot link constraints (Greedy Additive Edge Contraction) and they compared both with other greedy and local-search algorithms for solving the multicut/correlation clustering problem.
\item Dynamically introduced cannot-link constraints were also proposed by \cite{wolf2018mutex}, that... \TODO{brief MWS description as efficient implementation of this update rule}. In the Supplementary material we show that the algorithm they present can be seen as a special case of UGACA with absolute maximum update rule (Table \ref{tab:linkage-criteria} row 2 - \emph{absolute maximum}) and that adding or not cannot-link constraints does not change the final clustering.
\item On the other hand, the signed graph versions of UGACA with update rules \emph{arithmetic mean}, \emph{max} and \emph{min} (rows 3 to 5 in Table \ref{tab:linkage-criteria}) are new algorithms that were never mentioned in the literature, \UPDATE{especially those with cannot-link constraints}.
\item In this article, we test and run a comparison of the five types of update rule listed in Table \ref{tab:linkage-criteria} (i.e. \emph{sum}, \emph{absolute mac}, \emph{arithmetic mean}, \emph{max} and \emph{min}), since \UPDATE{they they are the most commonly used and the cheapest to compute}. Nevertheless, UGACA can be easily generalized to include even more complex rules. For instance, \cite{nunez2013machine} proposes a learned version of update rule where a neural network updates the edge weights depending on edge and node features; another really common choice is to introduce a weight regularization depending on the size of the linked nodes (\cite{felzenszwalb2004efficient} and \cite{kardoostsolving} (unpublished...??), Thorsten..); 
\cite{funke2018large} on the other hand accumulate statistics for each \UPDATE{boundary between clusters} by keeping a discrete histograms of the edge weights and then computes quantiles from it.
\end{itemize}

>>>>

\section{Unified Graph Agglomerative Clustering}
In this section we introduce the Unified Graph Agglomerative Clustering Algorithm (UGACA) that represents a simple generalized formalization of many agglomerative graph clustering algorithms. \UPDATE{It can be used to describe both unsigned clustering algorithms ingesting positive node similarities and signed clustering algorithms using attractive and repulsive cues.}\\
First, we define notation and the concept of cannot-link constraints. Then, we introduce the generalized algorithm and show \UPDATE{how it can be used to introduce several new agglomerative clustering algorithms for graphs with signed edge weights.}

@@

Given an initial clustering $\Pi$ where each node of the original graph $\mathcal{G}$ represents its own cluster, the algorithm starts by initializing a contracted graph $\mathcal{G}'$ (see Definition \ref{def:contr_graph}) that will represent the clustering $\Pi$ at each iteration of the agglomeration. The algorithm then sorts all edges $e\in E$ by their absolute weight $|W(e)|$ in descending order in a priority queue, so that the most attractive and the most repulsive edges are processed first. It iteratively pops one edge $e_{uv}$ from priority queue and, depending on the value of the priority $\tilde{w}$ associated with $e_{uv}$, does the following: if the edge is attractive and $\tilde{w}>0$, then merge the connected nodes and perform an edge contraction of $e_{uv}$ in $\mathcal{G}'$ (see Sec. \ref{sec:contr_graph_and_CLC} and Fig. \ref{fig:edge_contraction_and_contr_graph}), provided that $e_{uv}$ was not previously flagged as a cannot-link constraint; for every new pair of double edges in $\mathcal{G}'$ update \UPDATE{their priorities} according to one of the update rules listed in Table \ref{tab:linkage-criteria} and update their cannot-link relationships (see example in Fig. \ref{fig:algorithm_with_without_CLC}b from step 3 to 4). If, on the other hand, the edge $e_{uv}$ is repulsive and $\tilde{w}\leq 0$, then flag it as a cannot-link constraint if and only if the algorithm input-option \texttt{addCannotLink} is true and mutual exclusion relationships should be enforced.

In the Supplementary material we provide a more detailed description of the algorithm (Sec. \ref{sec:detailed_algorithm}) and we also comment on its computational complexity (Sec. \ref{sec:complexity}).

 In the special case of an unsigned graph %$\mathcal{G}(V,E,W:E\rightarrow \mathbb{R}^+)$ 
 presenting strictly positive similarities between nodes, the algorithm returns only one single cluster as an output (see Fig. \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}a}), \UPDATE{together with a merging tree representing the merging order and defining a hierarchy of clusters} \TODO{use def of merging process of contracted graph to define merging tree?}.

@@

\subsection{Update rules and associated algorithms}
\begin{itemize}
\item In this section we will list some of the algorithms that can be described using the generalized Algorithm \ref{main_alg} presented above and we will introduce some new ones.
\item In Table \ref{tab:linkage-criteria} we collected the update rules that we tested and compared in this article. Later we will mention more of them that were proposed in the literature.
\item In the special case of unsigned graphs (Table \ref{tab:linkage-criteria} column 1), the UGAC algorithm becomes simply a version of hierarchical clustering with different linkage criteria depending on the chosen update rule. It outputs one single cluster and it also defines a hierarchy of clusters during the merging process that can be associated to \UPDATE{a merging tree/  the hierarchical clustering dendrogram}... \TODO{Problem: sum does not trivially define an ultra-metric-distance...}  
\item In the literature there have been already several methods using cannot-link constraints for agglomerative clustering algorithms \TODO{repetition from related work}: for example, \cite{malmberg2011generalized} used initially fixed mutual exclusion relationships, whereas \cite{levinkov2017comparative} introduced them dynamically in an edge contraction algorithm for signed graphs, named Greedy Fixation (Table \ref{tab:linkage-criteria} row 1, \emph{sum}). At the same time, they also proposed a version of it without cannot link constraints (Greedy Additive Edge Contraction) and they compared both with other greedy and local-search algorithms for solving the multicut/correlation clustering problem.
\item Dynamically introduced cannot-link constraints were also proposed by \cite{wolf2018mutex}, that... \TODO{brief MWS description as efficient implementation of this update rule}. In the Supplementary material we show that the algorithm they present can be seen as a special case of UGACA with absolute maximum update rule (Table \ref{tab:linkage-criteria} row 2 - \emph{absolute maximum}) and that adding or not cannot-link constraints does not change the final clustering.
\item On the other hand, the signed graph versions of UGACA with update rules \emph{arithmetic mean}, \emph{max} and \emph{min} (rows 3 to 5 in Table \ref{tab:linkage-criteria}) are new algorithms that were never mentioned in the literature, \UPDATE{especially those with cannot-link constraints}.
\item In this article, we test and run a comparison of the five types of update rule listed in Table \ref{tab:linkage-criteria} (i.e. \emph{sum}, \emph{absolute mac}, \emph{arithmetic mean}, \emph{max} and \emph{min}), since \UPDATE{they they are the most commonly used and the cheapest to compute}. Nevertheless, UGACA can be easily generalized to include even more complex rules. For instance, \cite{nunez2013machine} proposes a learned version of update rule where a neural network updates the edge weights depending on edge and node features; another really common choice is to introduce a weight regularization depending on the size of the linked nodes (\cite{felzenszwalb2004efficient} and \cite{kardoostsolving} (unpublished...??), Thorsten..); 
\cite{funke2018large} on the other hand accumulate statistics for each \UPDATE{boundary between clusters} by keeping a discrete histograms of the edge weights and then computes quantiles from it.
\end{itemize}

### 2019-03-27 chapters/5_experiments_cityscapes.tex
"Finish" draft of noise exp. section
Minor changes

% !TEX root = ../agglo_clust_review.tex
\subsection{Results on Cityscapes}

\begin{table}
    \centering
    \begin{subtable}[t!]{0.5\textwidth}\centering
        \begin{tabular}{l M{0.25\textwidth} M{0.25\textwidth}}
        \toprule
        & \multicolumn{2}{c}{\thead{Add Cannot-Link Constraints:}} \\
        Update rule: & \multicolumn{1}{c}{\thead{\textsc{No}}} & \multicolumn{1}{c}{\thead{\textsc{Yes}}} \\ \midrule
         % & \multicolumn{1}{c}{\thead{$\beta$}}  & \thead{AP} & \multicolumn{1}{c}{\thead{$\beta$}} & \thead{AP} \\ \midrule\midrule
\textbf{Mean}& \textbf{0.343}  & 0.339  \\
GMIS \cite{liu2018affinity} & 0.341 & -  \\
Max &   0.243  &   0.325  \\
Abs. Max. \cite{wolf2018mutex}  & -  & 0.321 \\
Sum \cite{levinkov2017comparative} & 0.313  & 0.319  \\
Min &  0.000    & 0.000  \\
        \end{tabular}
        % \caption{Linkage criteria}
    \end{subtable} 
    \caption{AP scores on the cityscapes validation set for UGACA and different types of update rules. \TODO{PROBLEM: all apart from GMIS uses finetuned affinities... Unfair comparison?} }
    \label{tab:results_cityscapes}
\end{table}

>>>>

% !TEX root = ../agglo_clust_review.tex
\subsection{Results on Cityscapes}
In the experiments on the cityscapes dataset, we used the pipeline proposed by \cite{liu2018affinity} \UPDATE{that currently represents the method not using any external dataset like MS-COCO}...
\begin{table}
    \centering
    \begin{subtable}[t!]{0.5\textwidth}\centering
        \begin{tabular}{l M{0.25\textwidth} M{0.25\textwidth}}
        \toprule
        & \multicolumn{2}{c}{\thead{Add Cannot-Link Constraints:}} \\
        Update rule: & \multicolumn{1}{c}{\thead{\textsc{No}}} & \multicolumn{1}{c}{\thead{\textsc{Yes}}} \\ \midrule
         % & \multicolumn{1}{c}{\thead{$\beta$}}  & \thead{AP} & \multicolumn{1}{c}{\thead{$\beta$}} & \thead{AP} \\ \midrule\midrule
\textbf{Mean}& \textbf{0.343}  & 0.339  \\
GMIS \cite{liu2018affinity} & 0.341 & -  \\
Max &   0.243  &   0.325  \\
Abs. Max. \cite{wolf2018mutex}  & -  & 0.321 \\
Sum \cite{levinkov2017comparative} & 0.313  & 0.319  \\
Min &  0.000    & 0.000  \\
        \end{tabular}
        % \caption{Linkage criteria}
    \end{subtable} 
    \caption{AP scores on the cityscapes validation set for UGACA and different types of update rules. \TODO{PROBLEM: all apart from GMIS uses finetuned affinities} }
    \label{tab:results_cityscapes}
\end{table}

### 2019-03-27 chapters/5_experiments_noise.tex
Minor changes

\caption{Plot illustrating Adapted RAND scores achieved by UGACA and different update rules when noise is added to the edge weights... \TODO{Label which uses only local-neighbors and which uses long-range connections}}\label{fig:noise_plots}
\end{figure*}

\subsection{Testing the robustness of the algorithms}
 In this section we will quantitatively test the robustness of UGACA by adding noise to the edge weights of the graph and see how different choices of update rules compare.

 Among the update rules listed in Table \ref{tab:linkage-criteria}, we will focus only on \emph{sum}, \emph{arithmetic mean} and \emph{absolute maximum}, since they achieved the best scores in the results presented in Sec. \ref{sec:exp_first_comparison}. Before to present the results shown in Fig. \ref{fig:noise_merge}, we will first introduce the type of noise that was added to the edge weights.

@@

 We then decided to use Perlin noise\footnote{In our experiments, we used opensimplex noise, that is a more recent open-source version of Perlin noise (\UPDATE{REF or link?})}, one of the most common gradient noises used in procedural pattern generation. This type of noise generates spatial random patterns that are locally smooth but have large and diverse variations on bigger scales. It generates values $n(x)\in[0,1]$ that were mapped to $[-\infty, \infty]$ with the Logit function, $N(x)=\mathrm{Logit}(n(x))$, and then combined with the original CNN predictions $F(x)=\mathrm{Logit}(p(x))$ in the two different following ways:
\begin{equation}
% \tilde{F}(x;\theta)=\begin{cases}
% F(x;\theta)+\mathcal{K}\cdot\max\left(N(x),0\right) & \text{if merge-biased}\\
% F(x;\theta)+\mathcal{K}\cdot\min\left(N(x),0\right) & \text{if split-biased}
% \end{cases}
\tilde{F}_{\pm}(x;\mathcal{K})=F(x)\pm\big|\mathcal{K}\cdot\max\left(\pm N(x),0\right)\big|,
\end{equation}
In this equation, $\mathcal{K}\in \mathbb{R}^+$ is a positive factor representing the amount of added noise; $\tilde{F}_{+}(x;\mathcal{K})$ represents a merge-biased modified version of the CNN predictions $F(x)$, such that the probability for two pixels to be in the same cluster is increased only if $N(x)>0$ (see Fig. \hyperref[fig:noisy_affs]{\ref*{fig:noisy_affs}b}); on the other hand, $\tilde{F}_{-}(x;\mathcal{K})$ represents a split-biased prediction with decreased probabilities only when $N(x)<0$ (Fig. \hyperref[fig:noisy_affs]{\ref*{fig:noisy_affs}c}). In the Supplementary material we provide a detailed description of the parameters used for the noise generation.

With this strategy, we randomly perturbed the CNN predictions and the edge weights only in some parts of the image. The plots in Fig. \ref{fig:noise_plots} represent the scores achieved by UGACA with different update rules depending on the amount of noise $\mathcal{K}$ added to the edge weights. 
\begin{itemize}
\item The experiments show that running UGACA on a grid graph with long-range connections always lead to better segmentations/clusterings \TODO{require def. of long range graph and long-range probability}
\item The most robust version of update rule proved to be the arithmetic mean, that achieves similar scores even when a significantly noisy edge weights.
\item The \emph{absolute maximum} update rule proposed by \cite{wolf2018mutex} provides an efficient and fast option, but it completely fails when too much noise is added.
\item The \emph{sum} update rule version was the slowest option tested, but it was not as robust as the \emph{mean} version, probably due to its tendency to grow one cluster at the time (see Sec. \ref{sec:exp_first_comparison}). Comment about MC energy (plots in Suppl. Material?)
\item On the other hand, UGACA with mean update rule and cannot-link constraints never achieves good overall scores, as we have already shown in Sec. \ref{sec:exp_first_comparison} and \UPDATE{Figure...?}, because \UPDATE{it tends to introduce too many false splits}. 
%Nevertheless, these additional experiments show that it provides an \UPDATE{almost true over-segmentation of the ground truth segmentation even with strongly merge-biased predictions}.
\end{itemize}

>>>>

\caption{Plot illustrating Adapted RAND scores achieved by UGACA and different update rules when noise is added to the edge weights... Solid lines represent median values, whereas values between the 25th and the 75th percentile are shown in shade areas.    \TODO{Label which uses only local-neighbors and which uses long-range connections}}\label{fig:noise_plots}
\end{figure*}

\subsection{Testing the robustness of the algorithms}
 In this section we will quantitatively test the robustness of UGACA by adding noise to the edge weights of the graph and see how different choices of update rules compare.
 Among the update rules listed in Table \ref{tab:linkage-criteria}, we will only focus on \emph{sum}, \emph{arithmetic mean} and \emph{absolute maximum}, since they achieved the best scores in the results presented in Sec. \ref{sec:exp_first_comparison}. %Before to present the results shown in Fig. \ref{fig:noise_plots}, we will first introduce the type of noise that was added to the edge weights.

@@

 We then decided to use Perlin noise\footnote{In our experiments, we used opensimplex noise, that is a more recent open-source version of Perlin noise (\UPDATE{REF or link?})}, one of the most common gradient noises used in procedural pattern generation. This type of noise generates spatial random patterns that are locally smooth but have large and diverse variations on bigger scales. It generates values $n(x)\in[0,1]$ that were mapped to $[-\infty, \infty]$ with the Logit function, $N(x)=\mathrm{Logit}(n(x))$, and then combined with the original CNN predictions $F(x)=\mathrm{Logit}(p(x))$ in the two different following ways:
\begin{equation}
% \tilde{F}(x;\theta)=\begin{cases}
% F(x;\theta)+\mathcal{K}\cdot\max\left(N(x),0\right) & \text{if merge-biased}\\
% F(x;\theta)+\mathcal{K}\cdot\min\left(N(x),0\right) & \text{if split-biased}
% \end{cases}
\tilde{F}_{\pm}(x;\mathcal{K})=F(x)\pm\big|\mathcal{K}\cdot\max\left(\pm N(x),0\right)\big|,
\end{equation}
In this equation, $\mathcal{K}\in \mathbb{R}^+$ is a positive factor representing the amount of added noise; $\tilde{F}_{+}(x;\mathcal{K})$ represents a merge-biased modified version of the CNN predictions $F(x)$, such that the probability for two pixels to be in the same cluster is increased only if $N(x)>0$ (see Fig. \hyperref[fig:noisy_affs]{\ref*{fig:noisy_affs}b}); on the other hand, $\tilde{F}_{-}(x;\mathcal{K})$ represents a split-biased prediction with decreased probabilities only if $N(x)<0$ (Fig. \hyperref[fig:noisy_affs]{\ref*{fig:noisy_affs}c}). In the Supplementary material we provide a detailed description of the parameters used for the noise generation.

With this strategy, we randomly perturbed the CNN predictions and the edge weights only in some parts of the image. The plots in Fig. \ref{fig:noise_plots} represent the scores achieved by UGACA with different update rules depending on the amount of noise $\mathcal{K}$ added to the edge weights. 
\begin{itemize}
\item The experiments show that running UGACA on a grid graph with long-range connections always lead to better segmentations/clusterings \TODO{require def. of long range graph and long-range probability}
\item The most robust version of update rule proved to be the arithmetic mean, that achieves similar scores even with significantly noisy edge weights.
\item The \emph{absolute maximum} update rule proposed by \cite{wolf2018mutex} provides an efficient and fast option, but it completely fails when too much noise is added.
\item The \emph{sum} update rule version was the slowest option tested, but it was not as robust as the \emph{mean} version, probably due to its tendency to grow one cluster at the time (see Sec. \ref{sec:exp_first_comparison}). Comment about MC energy (plots in Suppl. Material?)
\item On the other hand, UGACA with mean update rule and cannot-link constraints never achieves good overall scores, as we have already shown in Sec. \ref{sec:exp_first_comparison} and \UPDATE{Figure...?}, because \UPDATE{it tends to introduce too many false splits}. 
%Nevertheless, these additional experiments show that it provides an \UPDATE{almost true over-segmentation of the ground truth segmentation even with strongly merge-biased predictions}.
\end{itemize}
